---
title: "Introduccion_Plum"
author: "Marco A. Aquino-Lopez"
date: "3/5/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introducción a Plum

### What is a likelihood

Imagen there is a true age-depth function $t(x)$. 

```{r}
t <- function(x)(15*x)
curve(t,from = 0 , to = 20)
```
Now we have sample at depths $5$, $10$ and $15$ with a machine that tell us the date with and error of $\pm 20$. This will result in  3 samples coming from normal distribution with means $75$, $150$ and $225$ respectively and standar deviation $20$.


## Normal Distribution

```{r}
datos <- rnorm(1,50,10)
plot(dnorm(x =  seq(0,100),mean = 50,sd = 15),type = 'l',ylab = '',xlab='',main = "Normal Distribution")
points(x= datos,y = 0,pch = 16,col = rgb(0,0,1,.5))
#Sys.sleep(2)
for (i in 1:30){
  points(x= rnorm(1,50,15),y = 0,pch = 16,col = rgb(0,0,1,.5) )
#  Sys.sleep(1)
}

```

```{r}
datos <- rnorm(1,50,10)
plot(dnorm(x =  seq(0,100),mean = 50,sd = 3),type = 'l',ylab = '',xlab='',main = "Normal Distribution")
points(x= datos,y = 0,pch = 16,col = rgb(0,0,1,.5))
#Sys.sleep(2)
for (i in 1:30){
  points(x= rnorm(1,50,3),y = 0,pch = 16,col = rgb(0,0,1,.5) )
#  Sys.sleep(1)
}
```


```{r}
depths <- c(5,10,15)
ages   <- rnorm(3,t(depths),20)
curve(t,from = 0 , to = 20)
points(depths,ages,col=rgb(0,0,1,.7),pch=18)
```

given that we only see the dates.
If we assume a linear model with intercept at 0

```{r}
curve(t,from = 0 , to = 20,col=rgb(0,0,0,.1) )
points(depths,ages,col=rgb(0,0,1,.7),pch=18)
abline(a= 0 ,b = 10.6 ,col=rgb(1,0,0,.5))
abline(a= 0 ,b = 20.9 ,col=rgb(1,0,0,.5))
abline(a= 0 ,b = 12.9 ,col=rgb(1,0,0,.5))
```
Well we know that these dates came from a normally distribution. So the likelihood tell us how likely this ages are given the parameters of the model.

the likelihood would look like this

```{r}
Like <- function (b){
  A <- (ages-(b*depths))**2
  exp(-.5 * A/4)  
}

curve(Like,from = 00, to =25,ylab = '',xlab = "b",)
```
```{r}
curve(Like,from = 12, to =15)
```


```{r}
model <- lm(ages ~ depths-1)
summary(model)
plx<-predict(model, se=T)

curve(t,from = 0 , to = 20,col=rgb(0,0,0,.1) )
points(depths,ages,col=rgb(0,0,1,.7),pch=18)
abline(model)
lines(depths,plx$fit - qt(0.975,plx$df)*plx$se, lty=2)
lines(depths,plx$fit + qt(0.975,plx$df)*plx$se, lty=2)
```


### logaritmic funcion

```{r}
logFun <- function(x,A=100)( 10*log(A/(100-x)) )
curve(logFun,from = 0,to =100,main="Log function", ylab = '',xlab='')
```

```{r}
logFun <- function(x,A=100)( 10*log(A/(100-x)) )
x=seq(99.999999999,100,length.out=100)
plot(x,logFun(x),type = 'l',main="Log function", ylab = '',xlab='' )
```



### Bacon model

```{r}
alphas <- rgamma(5,shape = 1.5,scale = 20/1.5)
memory <- rbeta(1,shape1 = 5,shape2 = 5)
bacon <- function (al=alphas,mem=memory){
  w <- mem
  m <- c(al[1])
  for (i in al[-1]){
    m <- c(m,w*tail(m,1) + (1-w)*i )
  }
  rev(m) 
}

print(alphas )
print(bacon() )

plot(1:length(alphas),cumsum(bacon()), type = 'l',main="Bacon model",xlab = "depth",ylab = "age")

```
```{r}
setwd("~")
library(rbacon)
Bacon(ask=F,suggest=F)
```
#############
#############
#############








